---
title: "instructions"
author: "TSAI -YI FAN"
date: "2018年4月16日"
output: html_document
---

knitr::opts_chunk$set(echo = TRUE)

Loading and preprocessing the data
================
- Load the data
- Process/transform the data into a format suitable for your analysis.

```{r echo = TRUE}
tbl <- read.csv(file = "activity.csv")
head(tbl)
 
```
What is mean total number of steps taken per day?
==============
- Calculate the total number of steps taken per day
```{r echo = TRUE}
daily_steps <- tapply(tbl$steps, tbl$date, sum)
head(daily_steps)
 
```
- Make a histogram of the total number of steps taken each day
```{r echo = TRUE}
hist(daily_steps,breaks = 30, xlab = "Total daily Steps", main = "Histogram of Total Steps by day")
 
```
- Calculate and report the mean and median of the total number of steps taken per day
```{r echo = TRUE}
median(daily_steps,na.rm = T)
mean(daily_steps,na.rm = T)
 
```
What is the average daily activity pattern?
================
- Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r echo = TRUE}
library(ggplot2)
library(dplyr)
tbl_interval <- tbl %>% na.omit() %>% group_by(interval) %>% summarize(mean_steps= mean(steps)) 
ggplot(tbl_interval, aes(x=interval, y=mean_steps))+ geom_line()
 
```
- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo = TRUE}
max_5_minute <- tbl_interval[which(tbl_interval$mean_steps == max(tbl_interval$mean_steps)),]
print(max_5_minute)
 
```
Imputing missing values
=======
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r echo = TRUE}
sum(is.na(tbl))
 
```
- Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r echo = TRUE}
tbl_mgr <- merge(tbl, tbl_interval, by="interval")
tbl_mgr$steps[is.na(tbl_mgr$steps)] <- tbl_mgr$mean_steps[is.na(tbl_mgr$steps)]
tbl_mgr <- tbl_mgr[order(tbl_mgr$date),]
tbl_mgr <- tbl_mgr[,-c(4)]
head(tbl_mgr)
 
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
===========
```{r echo = TRUE}
tbl_2 <- tbl_mgr[,c(2,1,3)]
head(tbl_2)
 
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
===============
- The histogram differs from the very first verison with missing values.
- The median value has been increased and equals to the mean value with the new dataset.
```{r echo = TRUE}
daily_steps2 <- tapply(tbl_2$steps, tbl_2$date, sum)
hist(daily_steps2,breaks = 30, xlab = "Total daily Steps", main = "Histogram of Total Steps by day with new dataset ")
median(daily_steps2)
mean(daily_steps2)
 
```
Are there differences in activity patterns between weekdays and weekends?
================
- Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r echo = TRUE}
tbl_2$whichday <- ifelse(weekdays(as.Date(tbl_2$date)) %in% c("Samstag", "Sonntag"), "weekend", "weekday")
tbl_2$whichday <- as.factor(tbl_2$whichday)
head(tbl_2)
 
```
Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday 
Makedays or weekend days (y-axis).
```{r echo = TRUE}
tbl_interval2 <- tbl_2 %>% group_by(interval, whichday) %>% summarize(mean_steps= mean(steps)) 
ggplot(tbl_interval2, aes(x=interval, y=mean_steps))+ 
  geom_line() +
  facet_grid(whichday ~.) +
  xlab("Interval") + 
  ylab("Mean Steps") +
  ggtitle("Comparison of Average Number of Steps seprated by weekday and weekend")
 
```


